Smoothing seismicity techniques applied to seismic sources characterization and probabilistic seismic hazard analysis in Brazil

### 2015-01-14 manuscript.tex
gji + plain subfigure + partial translation + nadaraya ref correction

\documentclass[draft,grl]{agutex}

@@

\usepackage{lineno}
\linenumbers*[1]

@@

% See how to enter figures and tables at the end of the article, after references.
\usepackage{float}
%\usepackage{caption}
\usepackage{subcaption}

%----------------------------------------------------------------------------------------
%	Glossaries
%----------------------------------------------------------------------------------------

@@

\altaffiltext{1}{Department of Geophysics,
Institute of Astronomy, Geophysics and Atmospherical Sciences,
University of São Paulo, São Paulo, Brazil.}

\altaffiltext{2}{School of Applied Math, Getúlio Vargas Foundation, Rio de Janeiro, Brazil.}

@@

A proposta de Arthur \citet{frankel_1995} foi usar uma largura de banda fixa,
nomeada distância de correlação $d_F$, e aplicar o estimador de Nadaraya-Watson (\ref{sec:nadaraya})
para suavizar o histograma 2D da sismicidade utilizando uma função de núcleo gaussiana:
\begin{equation}
	\ensuremath{
		\tilde{n}_j = \frac{ \sum_{i} n_i \,e^{ - \left(\frac{\gls{sym:dij}}{\gls{sym:dF}}\right)^2}}
						   { \sum_{i}     e^{ - \left(\frac{\gls{sym:dij}}{\gls{sym:dF}}\right)^2}},
	}
	\label{eq:ni}
\end{equation}
onde $\tilde{n}_j$ é a taxa de sismicidade (número de sismos com magnitude $m$ maior que a mínima magnitude
\gls{sym:Md} do catálogo) suavizada
na célula $j$, $n_i$ é o número de sismos em cada outra célula $i$ e
	\gls{sym:dij} é \glsdesc{sym:dij}.

@@

Já Gordon \citet{woo_1996} propôs avaliar a contribuição de uma função de núcleo
de cada sismo $i$ ocorrido em $\boldsymbol{r}_i$
na célula centrada em $\boldsymbol{r}$ que dependa também de sua
magnitude $m$:
\begin{equation}
	\ensuremath{
		\gls{sym:Rrm} = \sum_{i=1}^{N} \frac{ K(\gls{sym:r} - \gls{sym:ri}, m)}
											{T({\gls{sym:ri})}},
	}
	\label{eq:Rrm}
\end{equation}
onde $N$ é o número de tremores $i$ no catálogo
e $T(\gls{sym:ri})$ é o período em que todo sismo de magnitude acima de $m$ é completamente observado
em \gls{sym:ri}.

@@

\subsubsection{Bandwidth computation}

\begin{equation}
	\ensuremath{
%		h_i, d_i = \underset{d_i \ge \gls{sym:dk}, h_i \ge \gls{sym:hk}}{\argmin}
		h_i, d_i = \argmin_{\substack{h_i \ge \gls{sym:hk} \\
						              d_i \ge \gls{sym:dk}}
				           }
		\left[ s \left(h_i,d_i
			 		  \arrowvert
					  \gls{sym:k_cnn},\gls{sym:a_cnn}
			     \right)
			   := h_i + \gls{sym:a_cnn}d_i
	    \right],
	}
	\label{eq:helms_cnn}
\end{equation}
onde \gls{sym:k_cnn} é o \glsdesc{sym:k_cnn},
	 \gls{sym:a_cnn} é o \glsdesc{sym:a_cnn},
	 \gls{sym:dk} é o \glsdesc{sym:dk} e
	 \gls{sym:hk} é o \glsdesc{sym:hk}.

As larguras de bandas são escolhidas localmente, podendo ser menor, aumentando a resolução
onde há maior densidade de informação, ou crescer onde há escassez de informação e menor resolução.

@@

\begin{acknowledgments}
This work was partially supported by a grant from the Spanish Ministry of Science and Technology.
\end{acknowledgments}

@@

\begin{thebibliography}{}

\providecommand{\natexlab}[1]{#1}
\expandafter\ifx\csname urlstyle\endcsname\relax
  \providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}\else
  \providecommand{\doi}{doi:\discretionary{}{}{}\begingroup
  \urlstyle{rm}\Url}\fi

\bibitem[{\textit{Atkinson and Sloan}(1991)}]{AtkinsonSloan}
Atkinson, K., and I.~Sloan (1991), The numerical solution of first-kind
  logarithmic-kernel integral equations on smooth open arcs, \textit{Math.
  Comp.}, \textit{56}(193), 119--139.

\bibitem[{\textit{Colton and Kress}(1983)}]{ColtonKress1}
Colton, D., and R.~Kress (1983), \textit{Integral Equation Methods in
  Scattering Theory}, John Wiley, New York.

\bibitem[{\textit{Hsiao et~al.}(1991)\textit{Hsiao, Stephan, and
  Wendland}}]{StephanHsiao}
Hsiao, G.~C., E.~P. Stephan, and W.~L. Wendland (1991), On the {D}irichlet
  problem in elasticity for a domain exterior to an arc, \textit{J. Comput.
  Appl. Math.}, \textit{34}(1), 1--19.

\bibitem[{\textit{Lu and Ando}(2012)}]{LuAndo}
Lu, P., and M.~Ando (2012), Difference of scattering geometrical optics
  components and line integrals of currents in modified edge representation,
  \textit{Radio Sci.}, \textit{47},  RS3007, \doi{10.1029/2011RS004899}.

\end{thebibliography}

@@

\begin{figure}
	\centering
	\begin{subfigure}[b]{0.3\textwidth}
		\includegraphics[width=0.4\textwidth]{placeholder}
		\caption{caption A}
		\label{placeholder_a}
	\end{subfigure}
	\begin{subfigure}[b]{0.3\textwidth}
		\includegraphics[width=0.4\textwidth]{placeholder}
		\caption{caption B}
		\label{placeholder_b}
	\end{subfigure}
	\caption{main caption}
	\label{placeholder}
\end{figure}

>>>>

\documentclass[draft, grl]{agutex}

@@

% \usepackage{lineno}
% \linenumbers*[1]

@@

% See how to enter figures and tables at the end of the article, after references.
\usepackage[utf8]{inputenc}
\usepackage{float}
%\usepackage{caption}
%\usepackage{subcaption}

% \usepackage[round,sort,nonamebreak]{natbib} % citação bibliográfica textual(plainnat-ime.bst)
% \bibpunct{(}{)}{;}{a}{\hspace{-0.7ex},}{,} % estilo de citação. Veja alguns exemplos em http://merkel.zoneo.net/Latex/natbib.php

\usepackage{amsfonts}
\usepackage{textcomp}

%-----------------------------------------------------------------------------
%	colored references
%-----------------------------------------------------------------------------
% \usepackage[usenames,svgnames,dvipsnames]{xcolor}
\usepackage[a4paper,top=2.5cm,bottom=3.0cm,left=2.0cm,right=2.0cm]{geometry}
% \usepackage[colorlinks=true,citecolor=DarkGreen,linkcolor=NavyBlue,urlcolor=DarkRed,filecolor=Green,bookmarksopen=true]{hyperref}
% \usepackage[all]{hypcap}      % soluciona o problema com o hyperref
%-----------------------------------------------------------------------------
%	Glossaries
%-----------------------------------------------------------------------------

@@

\altaffiltext{1}{Department of Geophysics,
Institute of Astronomy, Geophysics and Atmospherical Sciences (IAG),
University of São Paulo (USP), São Paulo, Brazil.}

\altaffiltext{2}{Applied Math School (EMAp), Getúlio Vargas Foundation (FGV), Rio de Janeiro, Brazil.}

@@

Arthur \citet{frankel_1995} smoothing seismicity proposal consists originally in use a called correlation distance $d_F$ as the kernel \textit{fixed} bandwidth and next to apply the Nadaraya-Watson \citep{nadaraya_1964, watson_1964} estimator to smooth a 2D seismicity histogram using a gaussian kernel:
\begin{equation}
	\ensuremath{
		\tilde{n}_j = \frac{ \sum_{i} n_i \,e^{ - \left(\frac{\gls{sym:dij}}{\gls{sym:dF}}\right)^2}}
						   { \sum_{i}     e^{ - \left(\frac{\gls{sym:dij}}{\gls{sym:dF}}\right)^2}},
	}
	\label{eq:ni}
\end{equation}
where $\tilde{n}_j$ is the smoothed seismicity (number of earthquakes with magnitude $m$ above the minimum magnitude \gls{sym:Md} in the catalog) on cell $j$. $n_i$ is the earthquake counting in each other cell $i$ and \gls{sym:dij} is \glsdesc{sym:dij}.

@@

Gordon \citet{woo_1996} suggested to evaluate the contribution of each earthquake $i$, far $\boldsymbol{r}_i$ from its kernel, on the cell centered on $\boldsymbol{r}$ in dependence of their magnitude $m$:
\begin{equation}
	\ensuremath{
		\gls{sym:Rrm} = \sum_{i=1}^{N} \frac{ K(\gls{sym:r} - \gls{sym:ri}, m)}
											{T({\gls{sym:ri})}},
	}
	\label{eq:Rrm}
\end{equation}
where $N$ is the number of earthquakes $i$ on the catalogue
and $T(\gls{sym:ri})$ is the timeframe which all earthquakes with magnitude above $m$ is completely observed on \gls{sym:ri}.

Any kernel could be applied on that definition. In practice Woo used a \citet{kagan_knopoff_1980} for a infinity spatial domain:
\begin{equation}
	\ensuremath{
		K_{KJ}(\gls{sym:r}, m \arrowvert \gls{sym:aW}) =  \frac{  \gls{sym:aW}  -1}{\pi\gls{sym:hm}^2}
							\left[ 1 + \frac{\gls{sym:r}^2}{\gls{sym:hm}^2} \right]^{-\gls{sym:aW}},
	}
	\label{eq:k_kj}
\end{equation}
where \gls{sym:aW} is, accordingly \citet{verejones_1992}, \glsdesc{sym:aW}.

To compute the magnitude-dependent bandwidth function \gls{sym:hm}, Gordon Woo suggested the follow relation
\begin{equation}
	\ensuremath{
		h(m\arrowvert \gls{sym:a0}, \gls{sym:a1}) = \gls{sym:a0}e^{\gls{sym:a1}m},
	}
	\label{eq:hm}
\end{equation}
where $a_0$ and $a_1$ are computed by the regression of the mean nearest distance $h$ from each magnitude bin $m \pm \mathrm{d}m$.

@@

\subsubsection{Local bandwidth computation}

The method implemented to compute the space and time bandwidths for each earthquake was the Couppled-Nearest-Neighbour \citep{helmstetter_2012}:
\begin{equation}
	\ensuremath{
		h_i, d_i = \argmin_{\substack{h_i \ge \gls{sym:hk} \\
						              d_i \ge \gls{sym:dk}}
				           }
		\left[ s \left(h_i,d_i
			 		  \arrowvert
					  \gls{sym:k_cnn},\gls{sym:a_cnn}
			     \right)
			   := h_i + \gls{sym:a_cnn}d_i
	    \right],
	}
	\label{eq:helms_cnn}
\end{equation}
where \gls{sym:k_cnn} is the \glsdesc{sym:k_cnn},
	 \gls{sym:a_cnn} is a \glsdesc{sym:a_cnn},
	 \gls{sym:dk} is the \glsdesc{sym:dk} e
	 \gls{sym:hk} is the \glsdesc{sym:hk}.

The bandwidths are defined locally. Could be small in high earthquake density regions and higher where earthquakes are rarelly or regions with information lack.

@@

\begin{acknowledgments}
The author thanks specially to Marcelo Assumpção for the support and critical discussion as well all IAG-USP seismological team. It is also grateful to Stéphane Drouet and Vincent Guigues for all helpful discussions.
\end{acknowledgments}

@@

% ---------------------------------------------------------------------------- %
% Bibliografia
\bibliographystyle{BibTeX/agufull08} 	% citação bibliográfica textual
\bibliography{bib/bibliografia}  		% associado ao arquivo:

% \begin{thebibliography}{}

% \providecommand{\natexlab}[1]{#1}
% \expandafter\ifx\csname urlstyle\endcsname\relax
%   \providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}\else
%   \providecommand{\doi}{doi:\discretionary{}{}{}\begingroup
%   \urlstyle{rm}\Url}\fi

% \bibitem[{\textit{Atkinson and Sloan}(1991)}]{AtkinsonSloan}
% Atkinson, K., and I.~Sloan (1991), The numerical solution of first-kind
%   logarithmic-kernel integral equations on smooth open arcs, \textit{Math.
%   Comp.}, \textit{56}(193), 119--139.

% \bibitem[{\textit{Colton and Kress}(1983)}]{ColtonKress1}
% Colton, D., and R.~Kress (1983), \textit{Integral Equation Methods in
%   Scattering Theory}, John Wiley, New York.

% \bibitem[{\textit{Hsiao et~al.}(1991)\textit{Hsiao, Stephan, and
%   Wendland}}]{StephanHsiao}
% Hsiao, G.~C., E.~P. Stephan, and W.~L. Wendland (1991), On the {D}irichlet
%   problem in elasticity for a domain exterior to an arc, \textit{J. Comput.
%   Appl. Math.}, \textit{34}(1), 1--19.

% \bibitem[{\textit{Lu and Ando}(2012)}]{LuAndo}
% Lu, P., and M.~Ando (2012), Difference of scattering geometrical optics
%   components and line integrals of currents in modified edge representation,
%   \textit{Radio Sci.}, \textit{47},  RS3007, \doi{10.1029/2011RS004899}.

% \end{thebibliography}

@@

\begin{figure}
	\begin{center}
	\begin{table}[H]
		\begin{tabular}{ p{0.5\textwidth} p{0.5\textwidth} }
		\footnotesize(a)
		& 
		\footnotesize(b)
		\\		
		\includegraphics[width=0.45\textwidth]{placeholder} &
		\includegraphics[width=0.45\textwidth]{placeholder}
		\end{tabular}
	\end{table}
	\caption{This tabled figure shows the first (a) image and the 2nd one (b). Even with the same bla ba.}
	\label{placeholder}
	\end{center}
\end{figure}

\begin{figure}
	\begin{center}
	\begin{table}[H]
		\begin{tabular}{ p{0.33\textwidth} 
						 p{0.33\textwidth} 
						 p{0.33\textwidth} }
		\footnotesize(a) &
		\footnotesize(b) &
		\footnotesize(c)
		\\				
		\includegraphics[width=0.33\textwidth]{z_img_pga_frankel} &
		\includegraphics[width=0.33\textwidth]{z_img_pga_woo} &
		\includegraphics[width=0.33\textwidth]{z_img_pga_helmstetter}
		\end{tabular}
	\end{table}
	\caption{This tabled figure shows the first (a) image and the 2nd one (b). Even with the same bla ba.}
	\label{placeholder2}
	\end{center}
\end{figure}
